\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-08-31 smurf_paper.tex
mostly PCA section updates

\title[\scuba: map-making with SMURF]{\scuba: iterative map-making
with the Sub-Millimetre User Reduction Facility}

@@

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centered over 450 and 850\,\micron. The camera is
  scanned across the sky and records data at a rate of 200\,Hz. This
  data volume is much greater than previously existing submillimetre
  cameras, and represents a significant data anlaysis challenge. We
  describe the production of SCUBA-2 maps using the Sub-Millimetre
  User Reduction Facility (SMURF) in which we have adopted a fast,
  iterative approach to map-making that enables data reduction on
  single, modern, high-end desktop computers with execution times that
  are shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole); and
  (iii) demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17).
\end{abstract}

@@

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO, another
camera from the same generation. Generally speaking, such methods are
fast, although dependending on the science goals and noise properties
of the data, they may not achieve the best noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times. However, PCA cleaning is
not a good generic solution for producing maps of extended structures,
since such sources produce correlated signals amongst many detectors,
and are removed by this procedure.

The best map-making strategies for recovering information on all
angular scales are \emph{maximum likelihood} techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimizes the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.][]{wright1996}, for the more general map-making
problem involving many detectors, the problem is significantly
complicated both by the need to measure the cross power-spectra for
all unique pairs of bolometers, and then to perform the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to SCUBA-2 data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, and correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modeled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar noise may be calculated for all of the data points from
a given bolometer (since the noise at any instant in time is
uncorrelated with others, nor with the data from other bolometers),
greatly simplifying the measurement of noise properties and the
inversion step that is so complicated in the maximum-likelihood
methods. In principle, it should also be possible to recover a wider
dynamic range in angular scales than the simplest ``direct'' methods,
with an increase in calculation time that is merely linear in the
number of iterations needed, while using a comparable amount of
memory. A fast iterative method for handling chopped data from CMB
experiments was described in \citet{wright1996}, and later applied to
chopped SCUBA data by \citet{johnstone2000}, although in each case the
chopping renders the data essentially white without the need to model
additional noise components. More similar to the algorithm described
here was the iterative pipeline recipe for fitting and removing baseline drifts
in SCUBA scan-map data as the astronomical image estimate improved
\citet{1999ASPC..172..171J}. The two closest modern relatives of SMURF are the
Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

@@

In this section we summarize how \scuba\ bolometers detect temperature
variations (Section~\ref{sec:bolos}), describe \scuba\ scan strategies
(Section~\ref{sec:scan}), describe the differences between SCUBA-2
Shared Risk Observing data and those of the the fully commissioned
instrument (Section~\ref{sec:s2sro}), typical bolometers signals
(Section~\ref{sec:bolosignal}), the impact of magnetic field pickup
(Section~\ref{sec:magpickup}, and finally illustrate the use of
principal component analysis (Section~\ref{sec:pca}).

%--------------------------------------------------
\subsection{How the \scuba\ bolometers work}
\label{sec:bolos}
%--------------------------------------------------

@@

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ approximately 140 and 200\,mK at 850 and 450\,\micron,
respectively. The range about $T_\mathrm{C}$ in which the transition
from a superconductor to a normal material occurs is extremely small
(...typical value...?). The entire \scuba\ focal plane is held at a
\emph{base temperature} somewhat below $T_\mathrm{C}$. The TESs are
then biased up into the transition using two systems. First,
individual pixel heaters (using a current bias to apply the desired
Joule heating) are set to fixed values at the start of an observation
to compensate for most of the loading from the instrument and
atmosphere (in better weather, less compensation from the heaters is
required). Then, an electrical voltage bias is applied. As the
instrument observes, and there are tiny temperature fluctuations due
to astronomical sources and atmosphere, the impedence
changes. Thermoelectrical feedback helps to keep the TESs ``locked''
in the transition as this occurs; an increase in temperature increases
the impedence, and lowers the current and resulting heating
power. Changing magnetic fields resulting from these current changes
are then amplified using chains of superconducting quantum
interference devices (SQUIDs), and the larger output current is
finally digitized. Each detector has its own SQUID for the first-stage
of the amplification, but the remaining stages occur within a common
chain of SQUIDs for each of the 32 columns. All 40 rows are then
read-out in sequence at a row-visit frequency of about 12\,kHz. This
multiplexing scheme greatly reduces the number of readout and bias
lines to each detector, making such large-format bolometer array
designs feasible. Such a high sample rate is unnecessary to produce
maps, so the data are re-sampled to approximately 200\,Hz.  It should
also be noted that there is an additional 41st row of SQUID readouts
that are not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

The response of the output digitized current, $I$, with respect to a
change in the power $P$ incident on the bolometer, $dI/dP$ depends on
the efficiency of the thermal absorber, how well the TES is biased
into the transition, and the gain of the SQUID amplifiers. In
practice, the response of a detector can therefore vary somewhat from
observation to observation depending on how the arrays have been
tuned, and the observing conditions \citep[see][]{holland2012}. To
track these changes, at the start of each observation, the current is
ramped through the bolometer heaters such that $dI/dP$ can be
measured. We will refer to the measurements of $dI/dP$ across the
focal plane as the \emph{flatfield} for a given observation.

In general, bolometers have two predominant noise components: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers (largely independent from bolometer to
bolometer); and (ii) longer timescale drift resulting from slow
variations in the mean load (e.g. fridge, telescope, or atmospheric
changes), and also the readout electronics (often correlated amongst
many or all of the bolometers). As such, the spectrum corresponding to
a bolometer time-series exhibits a ``$1/f$'' shape below some knee
frequency, $f_\mathrm{knee}$, and is relatively flat at higher
frequencies. Other smaller effects may include line features in the
spectrum (common culprits are aliased high-frequency noise detected
during the multiplexed readout of the arrays as described in
Section~\ref{sec:bolos}, pickup from AC mains, or microphonic signals
induced by mechanical vibrations), and finally a roll-off towards the
Nyquist frequency caused by the anti-aliasing filter that precedes
digitisation.

The final conversion of power received by the detector into
astronomical flux units requires two more fundamental corrections: (i)
atmospheric extinction (due to atmospheric water vapour) must be
compensated for as a function of time; and (ii) the corrected power is
related to an astronomical flux density through the effective area of
the detector, the spectral response of the bolometers (with bandpasses
set by optical filters), the spectrum of the source, and the optical
efficiencies of the dish surface and other optical components in the
system -- a \emph{flux conversion factor} (FCF). Extinction is
measured, in most cases, using the JCMT line-of-sight water vapour
radiometer (WVM) which is calibrated using sky dips and the CSO
225\,GHz tipper. The FCF is determined empirically using measurements
of standard sources. For further details on these calibration steps,
see \citet{dempsey2012}.

%--------------------------------------------------
\subsection{Scan strategies}
\label{sec:scan}
%--------------------------------------------------

In order to observe astronomical sources, it is clearly advantageous
that the signals appear in the region of the spectrum that offers the
highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
is achieved by modulating the signal; the original SCUBA primarily
used the JCMT chopping secondary to place sources into the spectrum at
about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
ultimately limited to recovering information on angular scales smaller
than the chop throw \citep[up to several arcmin,][]{holland1999}.  The
primary approach to map-making with \scuba\ is instead to slew the
entire telescope at such a rate that angular scales of interest are
crossed on time scales shorter than $1/f_\mathrm{knee}$, as with other
earlier ground-based instruments that did not use a chopping secondary
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}. The
desire to scan rapidly has two basic constraints: (i) while the
telescope can physically move up to about 1800 and
3600\,arcsec\,s$^{-1}$ in azimuth and elevation, respectively, the
acceleration limit is about 600\,arcsec\,s$^{-2}$ (in azimuth, larger
in elevation), meaning that significant time is waisted on
``turn-arounds'' at large scan speeds; and (ii) bolometers are
designed with a thermal time constant that balances response with
sensitivity. Taking these constraints into consideration, \scuba\ was
ultimately designed to operate at a maximum scan speed of
600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
sample rate of 200\,Hz (although in practice it is typically in the
range 170--180\,Hz), which at that scan velocity results in a sample
every 3.0\,arcsec, or approximately one third of the 450\,\micron\
diffraction-limited full-width at half-maximum (FWHM) -- a typical
rule-of-thumb for adequately sampling a Gaussian point spread function
(PSF). Similarly, the \scuba\ bolometers are designed to have a time
constant of $\sim$1\,ms in order to avoid distorting PSF-sized sources
at this maximum scan rate.

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster
``boustrophedon'' pattern, i.e. a series of parallel passes across the
field. However, such a scan yields extremely poor performace for
structures that are transverse to the scan direction, since a long
time will have ellapsed between each parallel step, and the signal
will therefore be totally dominated by the $1/f$ noise. In addition,
other periodic noise features may appear in the data. With such a
regularly spaced scan strategy, these noise features can easily
conspire with the scan pattern to produce similarly periodic noise
structures in the map, or \emph{scan-synchronous} noise. It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible offer the best protection against large-scale noise
\citep[see discussion in][and references therein]{kovacs2008b}; such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
\scuba\ presently has three primary scan modes: PONG, which covers
rectangular areas larger than the array footprint in a closed pattern
that ``bounces'' off the edges at 45\,degree angles (and also usually
combined with a rotation through a number of fixed position angles to
create a ``rotating PONG'' with even better cross-linking); Lissajous,
in which sinusoidal functions describe the motion along orthogonal
axes (also to cover rectangular areas -- the PONG and Lissajous modes
are both related to similar scan modes developed at the Caltech
Submillimetre Observatory); and a ``constant velocity daisy'' for
small areas (of order the array footprint), in which the telescope
moves in a circle, whose centre also slowly traces out a circle. For a
more complete description of the \scuba\ observing modes, see
\citet{holland2012}.

%-------------------------------------------------
\subsection{Differences between the S2SRO and 2011 fully-commissioned instrument}
\label{sec:s2sro}
%-------------------------------------------------

The S2SRO period took place during February and March 2010. For the
purpose of this paper, we will consider data taken as early as 2009
December 3, at which point the instrument was being commissioned in a
similar configuration to that of the S2SRO period. At this time, each
of the 450 and 850\,\micron\ focal planes were populated with single
subarrays, s4a and s8d respectively. There were three main problems to
contend with during this period. First, the fridge was not able to
cool the focal plane to the desired base temperature (partly due to
the heat conductance of the wires from the arrays during this period)
which affected stability, particularly at 850\,\micron. Second, a
non-sinusoidal, large-scale oscillation in the fridge temperature
could easily be seen as a common signal to all of the detectors with a
period of approximately 25\,s.  Third, there were problems with the
subarray fabrication process which led to large numbers of dead
pixels, as well as large-scale variations in $T_\mathrm{C}$ and the
thermal conductivity, $G$. These latter issues made it difficult to
simulataneously place all of the TESs into the optimal region of the
superconducting transition, leading to significantly poorer
sensitivities. While there were moderate variations from
night-to-night during the observing period, there were typically 600
to 900 working detectors in each of the s4a and s8d subarrays.

Since S2SRO the fridge performance has been greatly improved using NTD
thermometry to servo the temperature to a constant value, and with new
wiring, adequate base temperatures can now be achieved. However, the
gradient in detector properties across the focal plane continues to be
a problem now that there is a full complement of subarrays (which
still includes the two subarrays used during S2SRO). Optimal tuning to
achieve the greatest overall array sensitivity is described in
\citet{holland2012}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{pspec.pdf}
\caption{Bolometer power spectral densities for the same two data sets
  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. Four
  of the most sensitive bolometers have been selected at each
  wavelength, and the flat-fielded (but otherwise raw) PSDs are shown
  as coloured dotted lines (the blue signals are for the same time
  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
  lines are the PSDs of the common-mode signals at each wavelength,
  and the solid coloured lines show the PSDs of the bolometers once
  the common-mode is removed.  Finally, the dashed black lines show
  the spectral shape produced by a point source given the scan speeds
  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
  reference, the top horizontal axes shows the conversion from
  frequency to angular scale. Horizontal dotted lines at $2 \times
  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
  respectively, are provided as a visual reference for the white-noise
  levels. In addition to $1/f$ and white noise components, and line
  features, all of the PSDs exhibit the gradual roll-off of the
  anti-aliasing filter just below the Nyquist frequency.  Left: for
  the S2SRO data there are broad line features in the PSDs at both
  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
  450 and 850\,\micron. Common-mode subtraction removes most of the
  correlated fridge oscillation signal, lowering the $1/f$ knees to
  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
  respectively. Right: data from the fully-commissioned instrument
  tend to have lower $1/f$ knees, and fewer line features. The
  remaining low-frequency noise, now that the fridge oscillations have
  been removed, is considerably more independent from bolometer to
  bolometer than in the S2SRO data (note the larger spread in the
  dotted lines at low frequencies compared to the left-hand plot,
  particularly at 450\,\micron). Finally, this example shows that the
  white noise performance is in factor similar for the two subarrays
  (s4a and s8b) that were used both before and after the upgrades.}
\label{fig:pspec}
\end{figure*}

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (note that variations are still seen in the mixing chamber; it
is likely that these are not observed at the focal plane itself, where
the NTD thermometry used in the servo loop is performed). There is
minimal correlated low-frequency noise in these 450\,\micron\ data
(note that there is almost no difference once the common-mode has been
subtarcted). The 850\,\micron\ channel, however, exhibits a more
significant signal that is obviously correlated with the telescope
motion. It is also correlated amongs many of the detectors, and
common-mode removal corrects it to some extent. Part of the reason
that this is seen here, and not in the S2SRO data set shown, is that
the amplitude of the scan pattern is larger. Most of this
``scan-synchronous'' noise is attributed to magnetic field pickup, as
described in Section~\ref{sec:magpickup}.

Next, we produce power spectral density (PSD) plots for four of the
most sensitive bolometers from both focal planes, using the same two
data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
the convention that the PSD as a function of frequency,
$\mathbf{P}(f)$, is normalized such that the integral over frequency
gives the same variance as the time-series variance across the full
time series. In other words, given a bolometer signal $\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
df ,
\end{equation}
%
where we only integrate over the positive frequencies up to
$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
for the (equal) power that appears at negative frequencies in the
discrete Fourier Transform. The units of the PSD written in this form
are pW$^2$\,Hz$^{-1}$; without the factor of 2 in front the convention
would be to write the units as pW$^2$\,s to distinguish the two
normalisations.

>>>>

\title[\scuba: iterative map-making with SMURF]{\scuba: iterative map-making
with the Sub-Millimetre User Reduction Facility}

@@

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centered over 450 and 850\,\micron. The camera is
  scanned across the sky and records data at a rate of 200\,Hz. This
  data volume is much greater than that of previously existing
  submillimetre cameras, and represents a significant data anlaysis
  challenge. We describe the production of SCUBA-2 maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products at the Canadian Astronomical Data Centre. A suite
  of tools are also provided for interactive data analysis.  Three
  detailed case studies are used to: (i) explore convergence
  properties of the map-maker using simple prior constraints (Uranus
  -- a point source); (ii) achieve the white-noise limit of the
  instrument for faint point-source studies (blank-field survey of the
  Lockman Hole); and (iii) demonstrate that our strategy is capable of
  recovering angular scales up to the size of the array footprint
  (approximately 5\,arcmin) for bright extended sources (star-forming
  region M17).
\end{abstract}

@@

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best noise performance on the angular scales
of interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are \emph{maximum likelihood} techniques, in which
the time-series data are expressed as a sampling of the ``true'' map
of the sky plus noise, and then the equation is inverted to estimate
the map as some weighted linear combination of the data that minimizes
the variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.][]{wright1996}, for the more general map-making
problem involving many detectors, the problem is significantly
complicated both by the need to measure the cross power-spectra for
all unique pairs of bolometers, and then to perform the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to SCUBA-2 data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, and correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modeled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar noise may be calculated for all of the data points from
a given bolometer (since the noise at any instant in time is
uncorrelated with others, nor with the data from other bolometers),
greatly simplifying the measurement of noise properties and the
inversion step that is so complicated in the maximum-likelihood
methods. In principle, it should also be possible to recover a wider
dynamic range in angular scales than the simplest ``direct'' methods,
with an increase in calculation time that is merely linear in the
number of iterations needed, while using a comparable amount of
memory. A fast iterative method for handling chopped data from CMB
experiments was described in \citet{wright1996}, and later applied to
chopped SCUBA data by \citet{johnstone2000}, although in each case the
chopping renders the data essentially white without the need to model
additional noise components. More similar to the algorithm described
here was the iterative pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The two closest modern
relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

@@

In this section we summarize how \scuba\ bolometers detect temperature
variations (Section~\ref{sec:bolos}), give exampels of typical
bolometers signals (Section~\ref{sec:bolosignal}), and the impact of
magnetic field pickup (Section~\ref{sec:magpickup}), and finally
illustrate the use of principal component analysis to explore the
correlated noise properties of \scuba\ data (Section~\ref{sec:pca}).

%--------------------------------------------------
\subsection{Description of \scuba}
\label{sec:bolos}
%--------------------------------------------------

@@

Incoming light passes first through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. The bolometers in each
focal plane consist of absorbers coupled to superconducting
transition-edge sensors (TESs) for thermometry. The focal planes are
held at a temperature somewhat below the critical temperatures of the
TESs by the fridge. The devices are then biased into the
superconducting transition using a combination of individual pixel
heaters (to compensate for most of the bulk variations in loading from
the atmosphere and other backgrounds), and an electrical voltage
bias. Throughout an observation the heater bias is fixed, but
smaller-scale load variations both from the backgrounds and, to a much
lesser extent, astronomical sources, significantly alter the TES
impedences as their temperatures vary. Thermoelectrical feedback then
keeps the TESs ``locked'' in the superconducting transition; an
increase in temperature increases the impedence, and lowers the
current, and therefore the heating resulting from the voltage
bias. These current changes produce varying magnetic fields that are
amplified using chains of superconducting quantum interference devices
(SQUIDs), and the larger output currents are digitized. Each detector
has its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each of the
32 columns. All 40 rows are read-out in sequence at a row-visit
frequency of about 12\,kHz. Such a high sample rate is unnecessary to
produce maps, so the data are re-sampled to approximately 200\,Hz
before writing to disk.  It should also be noted that there is an
additional 41st row of SQUID readouts that are not connected to
TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
common to each column's amplifier chain. The relationship between the
output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
established using \emph{flatfield} observations immediately prior to
science observations, in which the output signal is measured
throughout a ramp of the pixel heaters (which provide a known input
power); see Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012}. Finally, the conversion to astronomical flux
units from power involves a correction for atmospheric extinction
\citep[primarily using the JCMT Water Vapour Monitor to track
line-of-sight opacity variations, see Section~3 in][]{dempsey2012},
and the application of a flux conversion factor (FCF) which is
established from regular measurements of calibrators such as Uranus
\citep[Section~5 in][]{dempsey2012}.

The predominant noise features in \scuba\ bolometers are: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers (largely independent from bolometer to
bolometer); and (ii) longer timescale drifts resulting from slow
variations in the mean load (e.g. fridge, telescope, or atmospheric
changes), and also the readout electronics (often correlated amongst
many or all of the bolometers). As such, the power spectra of
bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
below some knee frequency, $f_\mathrm{knee}$, and they are then
relatively flat at higher frequencies. The spectra also include line
features which are predominantly produced by aliased high-frequency
noise sources detected during the multiplexed readout stage as
described above.  Finally, an anti-aliasing filter produces roll-off
in the spectrum at frequencies approaching Nyquist prior to the
200\,Hz re-sampling.

We note that the noise performace of \scuba\ has evolved over time. An
initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
during February and March 2010, during which each of the 450 and
850\,\micron\ focal planes were populated with single subarrays, s4a
and s8d respectively. In addition to having significantly fewer
available bolometers than the current fully-commissioned instrument
(and therefore a mapping speed reduced by a factor of $\sim$4),
instabilities in the fridge lead to a large periodic signal with a
period of about 25\,s that was correlated amongst the detectors. After
upgrading and commissioning the instrument, a servo using newly added
NTD thermometers in the focal planes has effectively mitigated this
problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
improvements to the magnetic shielding (reducing magnetic field pickup
as described in Section~\ref{sec:magpickup}), as well as more
effective removal of aliased noise sources (which has reduced both the
presence of line features and the mean white noise level).

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan \emph{patterns} are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e. a series of parallel passes across the field. However, such a
scan yields extremely poor performace for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or \emph{scan-synchronous} noise.  It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible, offer the best protection against large-scale noise
\citep[see discussion in][and references therein]{kovacs2008b}; such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
\scuba\ presently has two primary scan modes: PONG, which covers
rectangular areas larger than the array footprint in a closed pattern
that ``bounces'' off the edges at 45\,degree angles (and also usually
combined with a rotation through a number of fixed position angles to
create a ``rotating PONG'' with even better cross-linking) which was
based on a similar scan mode developed at the Caltech Submillimeter
Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
areas (of order the array footprint), in which the telescope moves in
a circle, whose centre also slowly traces out a circle. For a more
complete description of the \scuba\ observing modes, see Section~5 in
\citet{holland2012}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{pspec.pdf}
\caption{Bolometer power spectral densities for the same two data sets
  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
  noise slightly and clarify some features. Four of the most sensitive
  bolometers have been selected at each wavelength, and the
  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
  as coloured dotted lines (the blue signals are for the same time
  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
  lines are the PSDs of the common-mode signals at each wavelength,
  and the solid coloured lines show the PSDs of the bolometers once
  the common-mode is removed.  Finally, the dashed black lines show
  the spectral shape produced by a point source given the scan speeds
  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
  reference, the top horizontal axes shows the conversion from
  frequency to angular scale. Horizontal dotted lines at $2 \times
  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
  respectively, are provided as a visual reference for the white-noise
  levels. In addition to $1/f$ and white noise components, and line
  features, all of the PSDs exhibit the gradual roll-off of the
  anti-aliasing filter just below the Nyquist frequency.  Left: for
  the S2SRO data there are broad line features in the PSDs at both
  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
  450 and 850\,\micron. Common-mode subtraction removes most of the
  correlated fridge oscillation signal, lowering the $1/f$ knees to
  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
  respectively. Right: data from the fully-commissioned instrument
  tend to have lower $1/f$ knees, and fewer line features. The
  remaining low-frequency noise, now that the fridge oscillations have
  been removed, is considerably more independent from bolometer to
  bolometer than in the S2SRO data (note the larger spread in the
  dotted lines at low frequencies compared to the left-hand plot,
  particularly at 450\,\micron). Finally, this example shows that the
  white noise performance is in factor similar for the two subarrays
  (s4a and s8b) that were used both before and after the upgrades.}
\label{fig:pspec}
\end{figure*}

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (note that variations are still seen in the mixing chamber; it
is likely that these are not observed at the focal plane itself, where
the NTD thermometry used in the servo loop is performed). There is
minimal correlated low-frequency noise in these 450\,\micron\ data
(note that there is almost no difference once the common-mode has been
subtarcted). The 850\,\micron\ channel, however, exhibits a more
significant signal that is obviously correlated with the telescope
motion. It is also correlated amongst many of the detectors, and
common-mode removal corrects it to some extent. Part of the reason
that this is seen here, and not in the S2SRO data set shown, is that
the amplitude of the scan pattern is larger. Most of this
``scan-synchronous'' noise is attributed to magnetic field pickup, as
described in Section~\ref{sec:magpickup}.

Next, we produce power spectral density (PSD) plots for four of the
most sensitive bolometers from both focal planes, using the same two
data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
the convention that the PSD as a function of frequency,
$\mathbf{P}(f)$, is normalized such that the integral over frequency
gives the same variance as the time-series variance across the full
time series. In other words, given a bolometer signal $\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
df ,
\end{equation}
%
where we only integrate over the positive frequencies up to
$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
for the (equal) power that appears at negative frequencies in the
discrete Fourier Transform. The units of the PSD written in this form
are pW$^2$\,Hz$^{-1}$.
